%! TEX root = ../main.tex

\chapter{Derivations and Calculations}
\label{app:derivation}


\section{Approximate max. UV Irradiance of an EAS}
\label{app:cr-uv-irradiance}

We estimate the UV irradiance, $I_\text{UV}$, or Wattage deposited per area, 
from fluorescence light stemming from an extensive air shower. To illustrate 
that telescopes need to be incredibly sensitive to observe this phenomenon, we 
assume fantastic to unrealistically good conditions for the \UV light yield and 
related parameters. This results in an  upper limit to $I_\text{UV}$.

We consider an \EAS of energy $E_0 = \SI{e18}{\eV}$ and inclination 
$\uptheta=\SI{15}{\degree}$. We assume the behaviour of the particle cascade 
resulting from the deeply inelastic is completely determined by the 
Heitler-Matthews model (see \todo{write this} and 
\cite{matthewsHeitlerModelExtensive2005, rissePhotonAirShowers2006}). We arrive
at the value for the atmospheric depth (height), at which the shower reaches the
maximum multiplicity:

\begin{equation}
\label{eq:irradiance-multiplicity}
\Xmax \approx \SI{700}{\gram\per\centi\meter\squared}\qquad\left(\;\widehat{\approx}\;\SI{18.5}{\kilo\meter}\;\text{above earth surface}\footnotemark\right).
\end{equation}
\footnotetext{The calculated height above surface that corresponds to the given 
atmospheric depth varies a lot based on atmospheric variables. We have assumed a
purely isothermic atmosphere with $T=\SI{278}{\kelvin}$, and a standard density 
of $\rho=\SI{0.86}{\kilogram\per\meter\cubed}$ at an altitude of 
\SI{1400}{\meter} above sea level.}

At $\Xmax$, the EM component of the shower contains roughly
$1-\left(E_0\,/\,\xi^\pi_C\right)^{\beta-1}\approx92\%$ of the primary particle 
energy, with photons and electrons sharing the fraction to roughly equal parts.
Since only electrons contribute meaningfully to excitations of air molecules, it 
follows that $E_\text{UV}=0.92\cdot0.5\cdot\SI{e18}{\electronvolt}=
\SI{4.6e17}{\electronvolt}$ are available a priori to create fluorescence light.
We take an optimistic estimate of the fluorescence light yield 
$\text{FY}\approx8\gamma\,/\,\SI{}{\mega\electronvolt}$ from 
\cite{keilhauerNitrogenFluorescenceAir2013}. The majority of these photons stem 
from the 2P(0,0) transition of N$_2$ \cite{aveSpectrallyResolvedPressure2008}, 
which has a characteristic wavelength of \SI{337.1}{\nano\meter}, and an average 
radiation time of \SI{42}{\nano\second} 
\cite{leanContributionUltravioletIrradiance1989}. We assume that all available 
energy $E_\text{UV}$ is immediately converted to molecular excitations of N$_2$ 
at $\Xmax$, and then gradually released according to an exponential decay and 
recover an expression for the (UV) power output of the air shower.

\begin{equation}
\label{eq:irradiance-power}
P(t)= \text{FY}\,E_\text{UV}\,\frac{hc}{\SI{337.1}{\nano\meter}} \cdot\frac{e^{-\SI{42}{\nano\second}/t}}{t}   
\end{equation}

This results in a maximum power output of $P_\text{max}\approx\SI{19.4}{\watt}$. 
In a last step, we convert the maximum power output to an irradiance measured on
ground. For this, we must take into acount that the area illuminated by the 
\UV-light grows proportional to the distance squared to the source (location of 
$\Xmax$). Furthermore, \UV-light is attenuated by the atmosphere. More 
specifically, UV light with a wavelength of \SI{337}{\nano\meter} drops in 
intensity by about 13\%/km in clear conditions 
\cite[see Fig. 83 on page 103 of][]{baumAttenuationUltravioletLight2008}. Supposing that 
an observer is at least $d=\SI{18.5}{\kilo\meter}$ away from $\Xmax$ (compare 
\cref{eq:irradiance-multiplicity}), we estimate the total irradiance this observer 
measures as

\begin{equation}
\label{eq:irradiance}
I_\text{UV} = P_\text{max}\,\frac{e^{-0.14\,\frac{d}{\SI{}{\kilo\meter}}}}{4\pi\,d^2}\approx\SI{0.32}{\nano\watt\per\square\meter}.
\end{equation}
